---
title: "R Notebook"
output: html_notebook
---

Data is from IPUMS and was downloaded as .csv and saved in this repo

Load libraries
```{r}
library(tidyverse)
library(survey)
library(tmap)
library(rgdal)
library(sandwich)
library(lmtest)
```

Load data from IPUMS (https://usa.ipums.org/usa/) that is already restricted to only Kentucky
```{r}
df <- read_csv("usa_00019.csv")
```

Read in shapefile
```{r}
ky_shp <- readOGR("cb_2016_21_puma10_500k", layer = "cb_2016_21_puma10_500k", GDAL1_integer64_policy = TRUE) 
```

Filter to children 5-18 in Kentucky
```{r}
df_filtered <- df %>% filter(AGE < 19 & AGE > 4)
```

Creating new variables that are easier to interpret
```{r}
df_filtered <- df_filtered %>%
  mutate(
    poverty = ifelse(POVERTY < 100, 1, 0),
    no_int_access = ifelse(CINETHH == 3, 1, 0)
  )

df_filtered$race <- "All Others"
df_filtered$race[df_filtered$RACE == 1] <- "White"
df_filtered$race[df_filtered$RACE == 2] <- "Black"
df_filtered$race[df_filtered$HISPAN == 1] <- "Hispanic"
df_filtered$race[df_filtered$HISPAN == 2] <- "Hispanic"
df_filtered$race[df_filtered$HISPAN == 3] <- "Hispanic"
df_filtered$race[df_filtered$HISPAN == 4] <- "Hispanic"
```

write smaller dataset to .csv format
```{r}
write_csv(df_filtered, "ky_children_only.csv")
```

Survey analysis
```{r}
svy_df <- svydesign(ids = ~1, weights = ~PERWT, data = df_filtered)
ky_int <- svymean(~no_int_access, design = svy_df)
ky_pov_int <- svyby(~no_int_access, ~poverty, design = svy_df, svymean)
ky_puma <- svyby(~no_int_access, ~PUMA, design = svy_df, svymean)
ky_race <- svyby(~no_int_access, ~race, design = svy_df, svymean)
```

Print statewide results
```{r}
ky_int
ky_pov_int
ky_race
```

Linear regression model 
```{r}
#variables as factors for regression and setting White to be the race tested against for statistical signifance
df_filtered$puma_factor <- as.factor(df_filtered$PUMA)
df_filtered$race <- as.factor(df_filtered$race)
df_filtered$race <- relevel(df_filtered$race, "White")

#model
mod1 <- lm(no_int_access ~ POVERTY + race + puma_factor, weights = PERWT, data = df_filtered)

#heteroskedasticity robust standard errors
mod1$newse <- vcovHC(mod1, type="HC3")
mod1.display <- coeftest(mod1, mod1$newse)

#cleaning up the output a bit for display
tidy_output <- broom::tidy(mod1.display)
tidy_output[,2:5] <- round(tidy_output[,2:5], 3)
tidy_output <- knitr::kable(tidy_output)
tidy_output
```


Overall, 8% of children 5-18 in Kentucky don't have internet access at home. 16% of children in poverty don't have interent access, compared to 5% of children not in poverty. Children in poverty are 3x more likely not to have internet access at home. 

Write PUMA data to csv
```{r}
write_csv(ky_puma, "ky_puma_int_access.csv")
```


Prepare shapefile and merge in new data
```{r}
ky_shp@data$PUMA <- as.numeric(as.character(ky_shp@data$PUMACE10))
ky_shp@data <- full_join(ky_shp@data, ky_puma, by = "PUMA")
ky_shp@data$no_int_access <- ky_shp@data$no_int_access * 100

```

Create and save map
```{r}
png("ky_int_map.png", height = 250, width = 550)
tm_shape(ky_shp) +
  tm_polygons("no_int_access", style="quantile", title="Children Ages 5-18 Without Home Internet Access, 2016 \n(Percent)", palette = "Blues")
dev.off()
```
