---
title: "R Notebook"
output: html_notebook
---

Data is from IPUMS and was downloaded as .csv and saved in this repo

Load libraries
```{r}
library(tidyverse)
library(survey)
library(tmap)
library(rgdal)
```

Load data
```{r}
df <- read_csv("usa_00017.csv")
```


Read in shapefile
```{r}
ky_shp <- readOGR("cb_2016_21_puma10_500k", layer = "cb_2016_21_puma10_500k", GDAL1_integer64_policy = TRUE) 
```


Create list of Public Use Microdata Areas to keep in data
```{r}
PUMA_list <- as.numeric(as.character(ky_shp@data$PUMACE10))
```


Filter to children 5-18 in Kentucky
```{r}
df_filtered <- df %>% filter(AGE < 19 & AGE > 4 & PUMA %in% PUMA_list)
```


```{r}
df_filtered <- df_filtered %>%
  mutate(
    poverty = ifelse(POVERTY < 100, 1, 0),
    no_int_access = ifelse(CINETHH == 3, 1, 0)
  )
```


write smaller dataset to .csv format
```{r}
write_csv(df_filtered, "ky_children_only.csv")
```

Survey analysis
```{r}
svy_df <- svydesign(ids = ~1, weights = ~PERWT, data = df_filtered)
ky_int <- svymean(~no_int_access, design = svy_df)
ky_pov_int <- svyby(~no_int_access, ~poverty, design = svy_df, svymean)
ky_puma <- svyby(~no_int_access, ~PUMA, design = svy_df, svymean)
```

Print statewide results
```{r}
ky_int
ky_pov_int
```

Overall, 9% of children 5-18 in Kentucky don't have internet access at home. 18% of children in poverty don't have interent access, compared to 6% of children not in poverty. Children in poverty are 3x more likely not to have internet access at home. 

Prepare shapefile and merge in new data
```{r}
ky_shp@data$PUMA <- as.numeric(as.character(ky_shp@data$PUMACE10))
ky_shp@data <- full_join(ky_shp@data, ky_puma, by = "PUMA")
ky_shp@data$no_int_access <- ky_shp@data$no_int_access * 100

```

Create and save map
```{r}
png("ky_int_map.png", height = 250, width = 550)
tm_shape(ky_shp) +
  tm_polygons("no_int_access", style="quantile", title="Children Ages 5-18 Without Home Internet Access \n(Percent)", palette = "Blues")
dev.off()
```